---
title: "Raster Tutorial in R"
author: "Gregory H. Halverson"
date: "January 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Origins

This tutorial is an interpretation of the neon data skills tutorial at:
http://neondataskills.org/R/Raster-Data-In-R/

## Load Libraries

The basic libraries needed for working with rasters in R are `raster`, `sp`, and `rgdal`.

```{r load_libraries}
library(raster)
library(sp)
library(rgdal)
```

## Load Raster

```{r load_raster}
dem <- raster('20170112222539_1989219947.tif')
dem
```

## Min/Max Values

Use `setMinMax` to define the minimum and maximum values for the raster.

```{r min_max}
dem <- setMinMax(dem)
dem
```

Use `cellStats` to check the range of values in the raster.

```{r check_min}
cellStats(dem, min)
```

```{r check_max}
cellStats(dem, max)
```

```{r check_range}
cellStats(dem, range)
```

## Coordinate Reference System

Check the projection using `crs`.

```{r projection}
crs(dem)
```

## Extent 

Check the extent of the raster using `extent`.

```{r extent}
extent(dem)
```

## Plotting

Plot the raster using `plot`.

```{r plot}
plot(dem, col=terrain.colors(30))
```

Or using `image`.

```{r image}
image(dem, col=terrain.colors(30))
```

## Distribution

Check the distribution of values in the raster using `hist`.

```{r histogram, echo=FALSE}
hist(dem)
```

